---
title: 'EARLY LIFE STRESS IN MICE ALTERS GUT MICROBIOTA INDEPENDENT OF MATERNAL MICROBIOTA'
author: "Keri Kemp"
date: "11/20/2020"
output:
  html_document: default
  fig_crop: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Raw Data
The original sequencing reads are available for download via NCBI’s Sequence Read Archive under Bioproject PRJNA605008.

``````{r message=FALSE}

# Set working directory and file paths 
setwd("~/Box/Kemp_MSEW_2020/new_analysis/R_analysis")
base_dir <- '~/Box/Kemp_MSEW_2020/new_analysis/R_analysis'

#Load libraries
library(phyloseq)
library(genefilter)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(compositions)
library(zCompositions)
library(CoDaSeq)
library(vegan)
library(btools)
library(rcompanion)
library(lme4)
library(lmerTest)
library(knitr)
library(kableExtra)
```

## Import the data into phyloseq
```{r}
# Create phyloseq object from otu and taxonomy tables from dada2, along with the sample metadata.
otu <- read.table("_bin/silva_otu_table2.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("_bin/silva_taxa_table2.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("_bin/mapping_file2.txt",sep="\t",header=T,row.names=1)
fitGTR <- readRDS("_bin/fitGTR.rds")
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon),phy_tree(fitGTR$tree))
ps

# add a DNAStringSet object (to be accessed by the refseq function) to keep track of the ASV sequences
sequences <- Biostrings::DNAStringSet(taxa_names(ps))
names(sequences) <- taxa_names(ps)
ps <- merge_phyloseq(ps, sequences)
ps

# rename our ASVs to something more convenient in downstream analysis, while automatically retaining the corresponding unique sequence identifier
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))

# remove chloroplasts, mitochondria, and Eukaryota
ntaxa(ps) #1631
ps <- subset_taxa(ps, Kingdom !="NA")
ntaxa(ps) #1607; 24 removed
ps <- subset_taxa(ps, Family !="Mitochondria")
ntaxa(ps) #1578; 29 removed
ps <- subset_taxa(ps, Order !="Chloroplast")
ntaxa(ps) #1570; 8 removed

sum(sample_sums(ps)) #10235191 #10235177

# filter very low abundance ASVs
ntaxa(ps) #1576
ps2<-filter_taxa(ps, function(x) mean(x) >5, TRUE)
ntaxa(ps2) #370
sum(sample_sums(ps2)) #10196018

# mean, max and min of sample read counts
min(sample_sums(ps2)) 
mean(sample_sums(ps2))
max(sample_sums(ps2)) 

# add read count
reads_sample <- microbiome::readcount(ps2)
sample_data(ps2)$reads_sample <- reads_sample
sample_data(ps2)

# remove samples with less than 3,000 reads
ps2 =subset_samples(ps2, reads_sample >3000)
min <- min(sample_sums(ps2))
sample_data(ps2) 
```

## Now export otu and taxa tables from phyloseq for codaseq
```{r}
otu = as(otu_table(ps2), "matrix")
taxon = as(tax_table(ps2), "matrix")
metadata = as(sample_data(ps2), "matrix")
otu = as.data.frame(otu)
alex.in <- otu
taxon = as.data.frame(taxon)
samples = as.data.frame(metadata) 
```

### Imput zero values and conduct the Aitchison center-log-ratio (CLR) transformation.
```{r}
# count number of zeros in the table
sum(otu == 0) # 17110

# Replace 0 values with an estimate (because normalization is taking log, can't have 0) using the Bayesian-multiplicative replacement function cmultRepl in Zcompositions. Also transposing here, because we need samples as rows.
d.czm <- zCompositions::cmultRepl(t(otu), method="CZM", label=0)
# No. corrected values:  1375

# Perform the centred log-ratio, or CLR (Aitchison) transformation using CoDaSeq.  
# The output will have the ASVs as columns the 69 samples as ROWS
d.clr <- CoDaSeq::codaSeq.clr(d.czm) # No. corrected values:  15978 
dim(d.clr) # 370  80

# Distance matrix
dist.clr <- dist(t(d.clr))
dist_pairs <- spaa::dist2list(dist.clr)
atch_pairs <- as.data.frame(dist_pairs)
names(atch_pairs)[1] <- "DAM_sample"
names(atch_pairs)[2] <- "SampleID"
names(atch_pairs)[3] <- "Distance"

```

### Define the function: Make knitr table from Permanova results. https://rdrr.io/github/TBrach/MicrobiomeX/src/R/Functions.R
```{r definition, echo = FALSE}
get_permanova_table <- function(adonis) {
  kable_table <- adonis[[1]]
  kable_table[["Pr(>F)"]] <- format(kable_table[["Pr(>F)"]], digits = 3)
  options(knitr.kable.NA = '') 
  knitr::kable(kable_table,
               digits = 3,
               caption = "Permanova results") %>%
  kable_styling(full_width = FALSE)
}
```

```{r}
get_permanova_table
```

### PD2 PERMANOVA
```{r}
############ PD2 PERMANOVA ############
# Subset metadata to just PD2 samples
pd2_samples <- samples %>%
  dplyr::filter(PD == "PD2")

# Make a list of PD2 sample names
pd2_list <- pd2_samples %>%
  pull(SampleID)  

# Filter the d.clr object to just the 10 pd2 samples
d.clr_pd2 <- subset(d.clr, select=pd2_list)
dim(d.clr_pd2)

# Permanova between groups using Aitchison distance
dist.clr <- dist(t(d.clr_pd2))
perm<-adonis(dist.clr~TREATMENT, pd2_samples, strata=pd2_samples$MOM, permutations=9999)
get_permanova_table(perm)
```

### PD10 PERMANOVA
```{r}
############ PD10 PERMONOVA ############
# Subset metadata to just PD10 samples
pd10_samples <- samples %>%
  dplyr::filter(PD == "PD10")

# Make a list of PD10 sample names
pd10_list <- pd10_samples %>%
  pull(SampleID)  

# Filter the d.clr object to just the pd10 samples
d.clr_pd10 <- subset(d.clr, select=pd10_list)
dim(d.clr_pd10)

# Permanova between groups using Aitchison distance
dist.clr <- dist(t(d.clr_pd10))
perm<-adonis(dist.clr~TREATMENT, strata=pd10_samples$MOM, pd10_samples, permutations=9999)
get_permanova_table(perm)
```

```{r}
############ PERMDISP PD10 ############
# Use betadisper to test for differences in dispersion of treatment groups at PD28
# First define groups
groups <- pd10_samples[["TREATMENT"]]

# Calculate dispersion from centroid
bdisp_PD10 <- betadisper(dist.clr , groups)

# Extract distances 
as.character(bdisp_PD10$group)->TREATMENT
bdisp_PD10_df <- cbind.data.frame(TREATMENT,bdisp_PD28$distances) #make a dataframe
bdisp_PD10_df <- add_rownames(bdisp_PD10_df, "SampleID") #make row names a column with the heading "SampleID"
names(bdisp_PD10_df)[3]<-"Dispersion" #rename column

dispersion_pd10_summary <- Rmisc::summarySE(bdisp_PD10_df, measurevar="Dispersion", groupvars=c("TREATMENT"))

# Welch Two Sample t-test
t.test(bdisp_PD10_df$Dispersion ~ bdisp_PD10_df$TREATMENT) 
# Permutation test for homogeneity of multivariate dispersions
permutest(bdisp_PD10, pairwise = FALSE, permutations=9999)
```

### PD28 PERMANOVA
```{r}
############ PD28 PERMANOVA ############
# Subset metadata to just PD28 samples
pd28_samples <- samples %>%
  dplyr::filter(PD == "PD28")

# Make a list of PD28 sample names
pd28_list <- pd28_samples %>%
  pull(SampleID)  

# Filter the d.clr object to just the pd28 samples
d.clr_pd28 <- subset(d.clr, select=pd28_list)
dim(d.clr_pd28)

# Distance matrix
dist.clr <- dist(t(d.clr_pd28))
# Permanova between groups using Aitchison distance
perm<-adonis(dist.clr~TREATMENT, pd28_samples, strata=pd28_samples$MOM, permutations=9999)
get_permanova_table(perm)
```

```{r}
############ PERMDISP PD28 ############
# Use betadisper to test for differences in dispersion of treatment groups at PD28
# First define groups
groups <- pd28_samples[["TREATMENT"]]

# Calculate dispersion from centroid
bdisp_PD28 <- betadisper(dist.clr , groups)

# Extract distances 
as.character(bdisp_PD28$group)->TREATMENT
bdisp_PD28_df <- cbind.data.frame(TREATMENT,bdisp_PD28$distances) #make a dataframe
bdisp_PD28_df <- add_rownames(bdisp_PD28_df, "SampleID") #make row names a column with the heading "SampleID"
names(bdisp_PD28_df)[3]<-"Dispersion" #rename column

dispersion_pd28_summary <- Rmisc::summarySE(bdisp_PD28_df, measurevar="Dispersion", groupvars=c("TREATMENT"))
dispersion_pd28_summary

# Welch Two Sample t-test
t.test(bdisp_PD28_df$Dispersion ~ bdisp_PD28_df$TREATMENT) 
# Permutation test for homogeneity of multivariate dispersions
permutest(bdisp_PD28, pairwise = FALSE, permutations=9999)
```

### PD28 PCA
```{r, fig.align='center', out.width="60%"}
############ PD28 PCA ############
# Samples must be ROWs and features/ASVs as COLUMNS. Use the prcomp included in the R core statistics.
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(t(d.clr_pd28))
# calculate percent variance explained for the axis labels
pc1 <- round((d.pcx$sdev[1]^2/sum(d.pcx$sdev^2))*100,1)
pc2 <- round((d.pcx$sdev[2]^2/sum(d.pcx$sdev^2))*100,1)
xlab <- paste("PC1: ", pc1, "%", sep="")
ylab <- paste("PC2: ", pc2, "%", sep="")

#extract NMDS scores
scrs_PD28 <- scores(d.pcx)
scrs_PD28 <- as.data.frame(scrs_PD28) %>%
  dplyr::select(PC1,PC2)
#To facilitate computing centroids, add Treatment as a variable to the data frame of scores
scrs_PD28 <- cbind(as.data.frame(scrs_PD28), Treatment = pd28_samples$TREATMENT)
# Now compute the group centroids, which are the mean coordinate on each axis, groupwise:
cent_PD28 <- aggregate(cbind(PC1, PC2) ~ Treatment, data = scrs_PD28, FUN = mean)
# To draw the spider, we need to use geom_segment() which requires coordiates to draw 
# the segment from and to. Our to coordinates, the xend and yend aesthetics will be 
# the centroids. So we need to replicate the group centroid for each observation in 
# the group. This we facilitate by a left join via merge:
segs_PD28 <- merge(scrs_PD28, setNames(cent_PD28, c('Treatment','oPC1','oPC2')),
                   by = 'Treatment', sort = FALSE)

# set colors
cols = c("Control" = "black", "MSEW" = "red")

# "spider" plot PCA with ggplot2
pd28_pca<-ggplot(scrs_PD28, aes(x = PC1, y = PC2, colour = Treatment)) +
  geom_segment(data = segs_PD28,
               mapping = aes(xend = oPC1, yend = oPC2)) + # spiders
  geom_point(data = cent_PD28, size = 4, shape = 21, fill= "white", show.legend = F) +                    
  coord_equal() +
  scale_colour_manual(values = c("Control" = "black", "MSEW" = "red")) +
  geom_point(aes(color = Treatment), size = 2, shape = 19)+
  theme_bw()+
  labs(color="TREATMENT",x=xlab, y=ylab) +
  theme(aspect.ratio = 1, # set plot to square
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        plot.margin = unit(c(0, 0, 2, 0), "cm"),
        text = element_text(size=16),axis.text.x=element_text(size=14))
# Add 95% confidence interval
pd28_pca <- pd28_pca + 
  stat_ellipse(linetype = 2)
pd28_pca
```

### Dam PERMANOVA
```{r}
############ Dams PERMANOVA ############
# Subset metadata to just dam samples
dam_samples <- samples %>%
  dplyr::filter(PD == "Dam")

# Make a list of Dam sample names
dam_list <- dam_samples %>%
  pull(SampleID)  

# Filter the d.clr object to just the dam samples
d.clr_dams <- subset(d.clr, select=dam_list)
dim(d.clr_dams)

# Permanova between groups using Aitchison distance.
dist.clr <- dist(t(d.clr_dams))
perm<-adonis(dist.clr~TREATMENT, dam_samples,  strata=dam_samples$MOM, permutations=9999)
get_permanova_table(perm)
```

### Dam PCA
```{r, fig.align='center', out.width="60%"}
############ Dams PCA ############
# Samples must be ROWs and features/ASVs as COLUMNS. Use the prcomp included in the R core statistics.
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(t(d.clr_dams))
# calculate percent variance explained for the axis labels
pc1 <- round((d.pcx$sdev[1]^2/sum(d.pcx$sdev^2))*100,1)
pc2 <- round((d.pcx$sdev[2]^2/sum(d.pcx$sdev^2))*100,1)
xlab <- paste("PC1: ", pc1, "%", sep="")
ylab <- paste("PC2: ", pc2, "%", sep="")

#extract NMDS scores
scrs_dams <- scores(d.pcx)
scrs_dams <- as.data.frame(scrs_dams) %>%
  dplyr::select(PC1,PC2)
#To facilitate computing centroids, add Treatment as a variable to the data frame of scores
scrs_dams <- cbind(as.data.frame(scrs_dams), Treatment = dam_samples$TREATMENT)
# Now compute the group centroids, which are the mean coordinate on each axis, groupwise:
cent_dams <- aggregate(cbind(PC1, PC2) ~ Treatment, data = scrs_dams, FUN = mean)
# To draw the spider, we need to use geom_segment() which requires coordiates to draw 
# the segment from and to. Our to coordinates, the xend and yend aesthetics will be 
# the centroids. So we need to replicate the group centroid for each observation in 
# the group. This we facilitate by a left join via merge:
segs_dams <- merge(scrs_dams, setNames(cent_dams, c('Treatment','oPC1','oPC2')),
                   by = 'Treatment', sort = FALSE)

# set colors
cols = c("Control" = "grey", "MSEW" = "#A32CC4")

# "spider" plot PCA with ggplot2
dams_pca<-ggplot(scrs_dams, aes(x = PC1, y = PC2, colour = Treatment)) +
  geom_segment(data = segs_dams,
               mapping = aes(xend = oPC1, yend = oPC2)) + # spiders
  geom_point(data = cent_dams, size = 4, shape = 21, fill= "white", show.legend = F) +                    
  coord_equal() +
  scale_colour_manual(values = c("Control" = "darkgray", "MSEW" = "#A32CC4")) +
  geom_point(aes(color = Treatment), size = 2, shape = 19)+
  theme_bw()+
  labs(color="TREATMENT",x=xlab, y=ylab) +
  theme(aspect.ratio = 1, # set plot to square
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        plot.margin = unit(c(0, 0, 2, 0), "cm"),
        text = element_text(size=16),axis.text.x=element_text(size=14))
# Add 95% confidence interval
dams_pca <- dams_pca + 
  stat_ellipse(linetype = 2)
dams_pca
```

### Alpha Diversity
```{r}
######################################### 
############ Alpha diversity ############
######################################### 

# Rarefy to account for differences in sequencing depth among samples. Trim OTUs from the dataset that are no longer observed in any sample (have a count of zero in every sample).
set.seed(310)
ELS_R = rarefy_even_depth(ps2, sample.size = 3200, replace = FALSE, trimOTUs = TRUE)
# Species Richness (se.chao1 = standard error of mean)
Chao1<-estimate_richness(ELS_R, measures=c("Chao1"))
Chao1$SampleID <- rownames(Chao1)
Chao1$SampleID<-chartr(old = ".", new = "-", Chao1$SampleID)
# Faith's Phylogenetic Diversity (SR = observed species richness)
FaithsPD<-btools::estimate_pd(ELS_R)
colnames(FaithsPD)[1] <-"Faiths_PD"
FaithsPD$SampleID <- rownames(FaithsPD)
# merge tables
alpha_div <- full_join(Chao1,FaithsPD, by = "SampleID")
alpha_div <- alpha_div[,c(3,5,1,2,4)]
# add metadata
metadata <- get_variable(ELS_R)
alpha_div <- full_join(alpha_div,metadata, by = "SampleID")
# write out file
write.csv(alpha_div,"alpha_div_3200.csv")

# summarize for stats and plotting
alpha_div$MOM <- as.factor(alpha_div$MOM)
fpd <- Rmisc::summarySE(alpha_div, measurevar="Faiths_PD", groupvars=c("PD","TREATMENT"))
chao <- Rmisc::summarySE(alpha_div, measurevar="Chao1", groupvars=c("PD","TREATMENT"))

# Subset for stats and plotting
PD2_adiv  <- subset(alpha_div,PD == "PD2" )
PD10_adiv  <- subset(alpha_div,PD == "PD10" )
PD28_adiv  <- subset(alpha_div,PD == "PD28" )
Dam_adiv  <- subset(alpha_div,PD == "Dam" & MOM != "dam_14") # exclude dam that had only one litter for paired tests

# Dam 014 removed
fpd2 <- Rmisc::summarySE(Dam_adiv, measurevar="Faiths_PD", groupvars=c("PD","TREATMENT"))
chao2 <- Rmisc::summarySE(Dam_adiv, measurevar="Chao1", groupvars=c("PD","TREATMENT"))

# write out file
fpd2 <- rbind.data.frame(fpd2,fpd)
chao2 <- rbind.data.frame(chao2,chao)
write.csv(fpd2,"Faiths_PD_Summary.csv")
write.csv(chao2,"Chao1_Summary.csv")
```

### PD28 Alpha Diversity Plots
```{r}
# subset summary statistics to PD28 for plotting
PD28_fpd <- subset(fpd, PD == "PD28")
PD28_chao <- subset(chao, PD == "PD28")
# Set plotting theme & colors
theme_set(theme_bw())
cols = c("Control" = "black", "MSEW" = "red")

# plot PD28 Faiths PD
fpd28_p <- ggplot(PD28_adiv, aes(x=TREATMENT, y=Faiths_PD)) +
  geom_errorbar(data=PD28_fpd, mapping = aes(x = TREATMENT, y = Faiths_PD, 
              ymin = Faiths_PD - se, ymax = Faiths_PD + se), size=0.5, color="dimgrey", width=.5) +
  geom_crossbar(data=PD28_fpd, aes(ymin = Faiths_PD, ymax = Faiths_PD),
              size=0.4,col="dimgrey", width = .5) +
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.3, height=0), size = 2)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "Phylogenetic Diversity \n (Faith's PD)") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=16),axis.text.x=element_text(size=16)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(size=16,colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
fpd28_p  

# plot PD28 Chao1
chao28_p <- ggplot(PD28_adiv, aes(x=TREATMENT, y=Chao1)) +
  geom_errorbar(data=PD28_chao, mapping = aes(x = TREATMENT, y = Chao1, 
              ymin = Chao1 - se, ymax = Chao1 + se), size=0.5, color="dimgrey", width=.5) +
  geom_crossbar(data=PD28_chao, aes(ymin = Chao1, ymax = Chao1),
              size=0.4,col="dimgrey", width = .5) +
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.3, height=0), size = 2)+
  scale_y_continuous(limits=c(120,240), breaks = seq(12, 240, by = 20))+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "ASV Richness \n  (Chao 1 Estimator)") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=16),axis.text.x=element_text(size=16)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(size=16,colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
chao28_p 
# combine PD28 alpha diveristy plots
theme_set(theme_bw())
alpha28_p <- cowplot::plot_grid(chao28_p, fpd28_p, ncol = 2, align = "v")
alpha28_p

```

### Dam Alpha Diversity Plots
```{r}
# subset summary statistics to Dams for plotting
Dam_fpd <- subset(fpd, PD == "Dam")
Dam_chao <- subset(chao, PD == "Dam")
# Set plotting theme & colors
theme_set(theme_bw())
cols = c("Control" = "grey", "MSEW" = "#A32CC4")

# plot dam Faiths PD
fpdD_p <- ggplot(Dam_adiv, aes(x=TREATMENT, y=Faiths_PD)) +
  geom_errorbar(data=Dam_fpd, mapping = aes(x = TREATMENT, y = Faiths_PD, 
                                             ymin = Faiths_PD - se, ymax = Faiths_PD + se), size=0.5, color="dimgrey", width=.5) +
  geom_crossbar(data=Dam_fpd, aes(ymin = Faiths_PD, ymax = Faiths_PD),
                size=0.4,col="dimgrey", width = .5) +
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 2)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "Phylogenetic Diversity \n (Faith's PD)") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=16),axis.text.x=element_text(size=16)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(size=16,colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
fpdD_p  
# plot dam Chao1
chaoD_p <- ggplot(Dam_adiv, aes(x=TREATMENT, y=Chao1)) +
  geom_errorbar(data=Dam_chao, mapping = aes(x = TREATMENT, y = Chao1, 
                                              ymin = Chao1 - se, ymax = Chao1 + se), size=0.5, color="dimgrey", width=.5) +
  geom_crossbar(data=Dam_chao, aes(ymin = Chao1, ymax = Chao1),
                size=0.4,col="dimgrey", width = .5) +
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 2)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "ASV Richness \n (Chao 1 Estimator)") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=16),axis.text.x=element_text(size=16)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(size=16,colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
chaoD_p 

# combine dam alpha diveristy plots
theme_set(theme_bw())
alphaDam_p <- cowplot::plot_grid(chaoD_p, fpdD_p, ncol = 2, align = "v")
alphaDam_p
```

### Alpha Diversity Normality Tests
```{r}
# Histograms for Chao1
# Create 2x2 plot environment  
par(mfrow = c(3, 3))
plotNormalHistogram(PD2_adiv$Chao1, main= "PD2 Chao1")
plotNormalHistogram((PD10_adiv$Chao1),main= "PD10 Chao1")
plotNormalHistogram(log(PD10_adiv$Chao1),main= "PD10 Chao1")
plotNormalHistogram(PD28_adiv$Chao1, main= "PD28 Chao1")
plotNormalHistogram(Dam_adiv$Chao1, main= "Dam Chao1")

# Test for normality 
shapiro.test(PD2_adiv$Chao1)
shapiro.test(log(PD10_adiv$Chao1)) 
shapiro.test(PD28_adiv$Chao1) 
shapiro.test(Dam_adiv$Chao1) 

# Histograms for Phylogenetic Diveristy 
# Create 2x2 plot environment  
par(mfrow = c(2, 2))
plotNormalHistogram(log(PD2_adiv$Faiths_PD), main= "PD2 Faiths PD")
plotNormalHistogram(PD10_adiv$Faiths_PD, main= "PD10 Faiths PD")
plotNormalHistogram(PD28_adiv$Faiths_PD, main= "PD28 Faiths PD")
plotNormalHistogram(Dam_adiv$Faiths_PD, main= "Dam Faiths PD")

# Test for normality 
shapiro.test((PD2_adiv$Faiths_PD)) 
shapiro.test(log(PD2_adiv$Faiths_PD)) 
shapiro.test(PD10_adiv$Faiths_PD)
shapiro.test(PD28_adiv$Faiths_PD)
shapiro.test(Dam_adiv$Faiths_PD) 
# all normal
```

### GLM Dam Alpha Diversity
```{r}
############ Dam stats ############
############ Dam Chao1 ############

model = lmer(Dam_adiv$Chao1 ~ TREATMENT + (1|MOM) + (1|EXP), data=Dam_adiv)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Chao1 ANOVA table for dams") %>%
  kable_styling(full_width = FALSE) 

# Create plot environment
par(mfrow = c(3, 2))
# plot data
plotNormalHistogram(Dam_adiv$Chao1, main= "Dam Chao1")
boxplot(Chao1 ~ TREATMENT, data = Dam_adiv)
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x, main= "Residuals")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))

############ Dam Faith's PD ############

model = lmer(Dam_adiv$Faiths_PD ~ TREATMENT + (1|MOM) + (1|EXP), data=Dam_adiv)
summary(model)
sstable <-anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Faith's PD ANOVA table for dams") %>%
  kable_styling(full_width = FALSE) 

# Create plot environment
par(mfrow = c(3, 2))
# plot data
plotNormalHistogram(Dam_adiv$Faiths_PD, main= "Dam Faith's PD")
boxplot(Faiths_PD ~ TREATMENT, data = Dam_adiv)
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x, main= "Residuals")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

### GLM PD28 Alpha Diversity
```{r}
############ PD28 stats ############
############ PD28 Chao1 ############
model = lmer(PD28_adiv$Chao1 ~ TREATMENT * (1|MOM) , data=PD28_adiv)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Chao1 ANOVA table for PD28") %>%
  kable_styling(full_width = FALSE) 

# Create plot environment
par(mfrow = c(3, 2))
# plot data
plotNormalHistogram(PD28_adiv$Chao1, main="Chao1")
boxplot(Chao1 ~ TREATMENT, data = PD28_adiv,
        ylab="Chao1",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x,main= "Residuals")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))


############ PD28 Faith's PD ############
model = lmer(PD28_adiv$Faiths_PD ~ TREATMENT + (1|MOM), data=PD28_adiv)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Faith's PD ANOVA table for PD28") %>%
  kable_styling(full_width = FALSE) 

# Create plot environment
par(mfrow = c(3, 2))
# plot data
plotNormalHistogram(PD28_adiv$Faiths_PD, main= "PD28 Faiths PD")
boxplot(Faiths_PD ~ TREATMENT, data = PD28_adiv,
        ylab="Faiths_PD",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x, main= "Residuals")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

### GLM PD10 Alpha Diversity
```{r}
############ PD10stats #############
############ PD10 Chao1 ############
model = lmer(log(PD10_adiv$Chao1) ~ TREATMENT + (1|MOM), data=PD10_adiv)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Chao1 ANOVA table for PD10") %>%
  kable_styling(full_width = FALSE) 

# Create 2x2 plot environment
par(mfrow = c(3, 2))
# plot data
plotNormalHistogram(log(PD10_adiv$Chao1), main= "PD10 Chao1")
boxplot(log(Chao1) ~ TREATMENT, data = PD10_adiv,
        ylab="Chao1",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x,  main= "Residuals")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))

############ PD10 Faith's PD ############
model = lmer(PD10_adiv$Faiths_PD ~ TREATMENT + (1|MOM), data=PD10_adiv)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Faith's PD table for PD10") %>%
  kable_styling(full_width = FALSE) 

# Create plot environment
par(mfrow = c(3, 2))
# plot data
plotNormalHistogram(log(PD10_adiv$Faiths_PD),main= "PD10 Faiths_PD")
boxplot(Faiths_PD ~ TREATMENT, data = PD10_adiv,
        ylab="Faiths_PD",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x, main="Residuals")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

### GLM PD2 Alpha Diversity
```{r}
############ PD2 stats ############
############ PD2 Chao1 ############
model = lmer(PD2_adiv$Chao1 ~ TREATMENT + (1|MOM), data=PD2_adiv)
summary(model)
sstable <-anova(model)
sstable

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Chao1 ANOVA table for PD2") %>%
  kable_styling(full_width = FALSE) 

# Create plot environment
par(mfrow = c(3, 2))
plotNormalHistogram(PD2_adiv$Chao1 , main= "PD2 Chao1")
# plot data
boxplot(Chao1 ~ TREATMENT, data = PD2_adiv,
        ylab="Log trans. Chao1",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x, main= "Residuals")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))

############ PD2 Faith's PD ############
model = lmer(log(PD2_adiv$Faiths_PD) ~ TREATMENT + (1|MOM), data=PD2_adiv)
summary(model)
sstable <-anova(model, type=1)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Faith's PD ANOVA table for PD2") %>%
  kable_styling(full_width = FALSE)

# Create plot environment
par(mfrow = c(3, 2))
# plot data
plotNormalHistogram(log(PD2_adiv$Faiths_PD), main= "PD2 Faith's PD")
boxplot(Faiths_PD ~ TREATMENT, data = PD2_adiv,
        ylab="Faiths_PD",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x, main= "Residules")
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

### Shared OTUs
```{r}
##################################################
############ Shared OTUs with mothers ############
##################################################
# normalize to relative abundance based on rarefied data
ELSnorm <- transform_sample_counts(ps2, function(x) x/sum(x ))
# now extract the transformed OTU table into a dataframe for vizualizations
ELSnorm_otu<-as.data.frame(t(otu_table(ELSnorm)))
ELSnorm_otu$OTU<-row.names(ELSnorm_otu)

# metadata
metadata = as(sample_data(ELSnorm), "matrix")
metadata = as.data.frame(metadata)
```

### PD28 Shared OTUs
```{r}
############ PD28 ############
metadata_PD28 <- subset(metadata, PD == "PD28")
# Create two vecotrs showing the variables you are interested in
interest1 <- as.character(metadata_PD28$SampleID)
interest2 <- as.character(metadata_PD28$DAM_sample)

# Design a function to do the compare
compare <- function(int1, int2, dat){
  dat2 <- dat %>% 
    dplyr::select(OTU, int1, int2) %>%
    dplyr::filter(!!sym(int1) > 0 , !!sym(int2) > 0)%>%
    dplyr::select(-int2)
  return(dat2)
}

# Loop through interest1 and interest2
shared_otus <- map2(interest1, interest2, compare, dat = ELSnorm_otu) %>%
  purrr::reduce(full_join, by = "OTU")
write.csv(shared_otus,"PD28_shared_ASVs_table.csv",row.names=FALSE)

# get the count how many OTUs each pup shares with its mom
sum_otu_counts <- as.data.frame(t(shared_otus %>%
                                    dplyr::select(-c(`OTU`)) %>%  
                                    summarise_all(funs(sum(!is.na(.))))))
names(sum_otu_counts)[1] <- "Count"

# get the total relative abundance of the OTUs each pup shares with its mom
sum_otu_abund <- as.data.frame(shared_otus %>%
                                 dplyr::select(-c(`OTU`)) %>%  
                                 colSums(na.rm = TRUE)) 
names(sum_otu_abund)[1] <- "Tot.Rel.Abund."

# get metadata to attach to the summary tables
variables <- as.data.frame(metadata %>%
                             dplyr::filter(PD == "PD28") %>%
                             dplyr::select(SampleID, PD, EXP, TREATMENT, LITTER, MOM, DAM_sample))

# combine summary tables and attach metadata 
PD28_shared_otus <- merge(as.data.frame(sum_otu_abund), as.data.frame(sum_otu_counts), by='row.names', all=TRUE)
"SampleID" -> colnames(PD28_shared_otus)[1]
PD28_shared_otus <- full_join(PD28_shared_otus,variables, by = "SampleID")
PD28_shared_otus <- inner_join(atch_pairs,PD28_shared_otus, by = c("SampleID","DAM_sample"))
# write out file
write.csv(PD28_shared_otus,"PD28_shared_ASVs.csv",row.names=FALSE)
```

### PD10 Shared OTUs
```{r}
############ PD10 ############
metadata_PD10 <- subset(metadata, PD == "PD10")
# Create two vecotrs showing the variables you are interested in
interest1 <- as.character(metadata_PD10$SampleID)
interest2 <- as.character(metadata_PD10$DAM_sample)

# Loop through interest1 and interest2
shared_otus <- map2(interest1, interest2, compare, dat = ELSnorm_otu) %>%
  purrr::reduce(full_join, by = "OTU")

# get the count how many OTUs each pup shares with its mom
sum_otu_counts <- as.data.frame(t(shared_otus %>%
                                    dplyr::select(-c(`OTU`)) %>%  
                                    summarise_all(funs(sum(!is.na(.))))))
names(sum_otu_counts)[1] <- "Count"

# get the total relative abundance of the OTUs each pup shares with its mom
sum_otu_abund <- as.data.frame(shared_otus %>%
                                 dplyr::select(-c(`OTU`)) %>%  
                                 colSums(na.rm = TRUE)) 
names(sum_otu_abund)[1] <- "Tot.Rel.Abund."

# get metadata to attach to the summary tables
variables <- as.data.frame(metadata %>%
                             dplyr::filter(PD == "PD10") %>%
                             dplyr::select(SampleID, PD, EXP, TREATMENT, LITTER, MOM, DAM_sample))
                               

# combine summary tables and attach metadata 
PD10_shared_otus <- merge(as.data.frame(sum_otu_abund), as.data.frame(sum_otu_counts), by='row.names', all=TRUE)
"SampleID" -> colnames(PD10_shared_otus)[1]
PD10_shared_otus <- full_join(PD10_shared_otus,variables, by = "SampleID")
PD10_shared_otus <- inner_join(atch_pairs,PD10_shared_otus, by = c("SampleID","DAM_sample"))
# write out file
write.csv(PD10_shared_otus,"PD10_shared_ASVs.csv", row.names=FALSE)
```

### Normality Tests for Shared OTUs
```{r}
################# normality and transformations #################
# https://rcompanion.org/handbook/I_12.html
# get metadata
metadata = as(sample_data(ELSnorm), "matrix")
metadata = as.data.frame(metadata)
metadata$MOM = as.factor(metadata$MOM)
metadata = subset(metadata, PD != "PD2")

# summarize for stats and plotting
sum_PD28RA <- Rmisc::summarySE(PD28_shared_otus, measurevar="Tot.Rel.Abund.", groupvars=c("TREATMENT"))
sum_PD28count <- Rmisc::summarySE(PD28_shared_otus, measurevar="Count", groupvars=c("TREATMENT"))
sum_PD28atch <- Rmisc::summarySE(PD28_shared_otus, measurevar="Distance", groupvars=c("TREATMENT"))
sum_PD10RA <- Rmisc::summarySE(PD10_shared_otus, measurevar="Tot.Rel.Abund.", groupvars=c("TREATMENT"))
sum_PD10count <- Rmisc::summarySE(PD10_shared_otus, measurevar="Count", groupvars=c("TREATMENT"))
sum_PD10atch <- Rmisc::summarySE(PD10_shared_otus, measurevar="Distance", groupvars=c("TREATMENT"))
# write out files
write.csv(sum_PD28RA,"Shared_ASVs_PD28_RA_Summary.csv")
write.csv(sum_PD28count,"Shared_ASVs_PD28_Count_Summary.csv")
write.csv(sum_PD28atch,"Shared_ASVs_PD28_AtchDist_Summary.csv")
write.csv(sum_PD10RA,"Shared_ASVs_PD10_RA_Summary.csv")
write.csv(sum_PD10count,"Shared_ASVs_PD10_Count_Summary.csv")
write.csv(sum_PD10atch,"Shared_ASVs_PD10_AtchDist_Summary.csv")

# transform proportional data with the Arcsine transformation (http://strata.uga.edu/8370/rtips/proportions.html)
PD10_shared_otus$PD10_RA_logit <-car::logit(PD10_shared_otus$Tot.Rel.Abund.)
PD28_shared_otus$PD28_RA_logit <-car::logit(PD28_shared_otus$Tot.Rel.Abund.)

par(mfrow = c(1, 2))
plotNormalHistogram((PD10_shared_otus$Count), main= "PD10 Count")
plotNormalHistogram((PD28_shared_otus$Count), main= "PD28 Count")

par(mfrow = c(1, 2))
plotNormalHistogram(PD10_shared_otus$PD10_RA_logit, main= "PD10 Logit Trans. Tot.Rel.Abund.")
plotNormalHistogram(PD28_shared_otus$PD28_RA_logit, main= "PD28 Logit Trans. Tot.Rel.Abund.")
# Logit looks better, use logit transformation

par(mfrow = c(1, 2))
plotNormalHistogram((PD10_shared_otus$Distance), main= "PD10 AtchDist")
plotNormalHistogram(PD28_shared_otus$Distance, main= "PD28 AtchDist")

# Test for normality 
shapiro.test(PD28_shared_otus$PD28_RA_logit) 
shapiro.test(PD10_shared_otus$PD10_RA_logit) 
shapiro.test((PD28_shared_otus$Count))
shapiro.test((PD10_shared_otus$Count))
shapiro.test(PD28_shared_otus$Distance)
shapiro.test(PD10_shared_otus$Distance)
# all normal
```

### GLM PD10 Shared OTUs
```{r}
################# PD10 stats #################
################# PD10 Atchison Dist ##################
model = lmer((PD10_shared_otus$Distance) ~ TREATMENT + (1|MOM), data=PD10_shared_otus)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Aitchison Distance ANOVA table for PD10") %>%
  kable_styling(full_width = FALSE) 

# Create 2x2 plot environment
par(mfrow = c(2, 2))
# plot data
boxplot((Distance) ~ TREATMENT, data = PD10_shared_otus,
        ylab="PD10",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x)
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

```{r}
################# PD10 RA #################
model = lmer(PD10_shared_otus$PD10_RA_logit ~ TREATMENT + (1|MOM), data=PD10_shared_otus)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 4, caption = "Relative Abundance ANOVA table for PD10") %>%
  kable_styling(full_width = FALSE) 

# Create 2x2 plot environment
par(mfrow = c(2, 2))
# plot data
boxplot(PD10_RA_logit ~ TREATMENT, data = PD10_shared_otus,
        ylab="PD10 Logit Trans. Tot. Rel. Abund.",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x)
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))

```

```{r}
################# PD10 Count #################
model = lmer((PD10_shared_otus$Count) ~ TREATMENT + (1|MOM), data=PD10_shared_otus)
summary(model)
sstable <- anova(model)


# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Count ANOVA table for PD10") %>%
  kable_styling(full_width = FALSE)

# Create 2x2 plot environment
par(mfrow = c(2, 2))
# plot data
boxplot(Count ~ TREATMENT, data = PD10_shared_otus,
        ylab="PD10 Count",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x)
# plot Q plot
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))

```

### GLM PD28 Shared OTUs
```{r}
################# PD28 stats #################
################# PD28 Atch Dist ##################
model = lmer(PD28_shared_otus$Distance ~ TREATMENT + (1|MOM), data=PD28_shared_otus)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Aitchison Distance ANOVA table for PD28") %>%
  kable_styling(full_width = FALSE)


# Create 2x2 plot environment
par(mfrow = c(2, 2))
# plot data
boxplot(Distance ~ TREATMENT, data = PD28_shared_otus,
        ylab="PD28 Aitchison",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x)
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

```{r}
################# PD28 RA #################
model = lmer(PD28_shared_otus$PD28_RA_logit ~ TREATMENT + (1|MOM), data=PD28_shared_otus)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 3, caption = "Relative Abundance ANOVA table for PD28") %>%
  kable_styling(full_width = FALSE)

# Create 2x2 plot environment
par(mfrow = c(2, 3))
# plot data
boxplot(PD28_RA_logit ~ TREATMENT, data = PD28_shared_otus,
        ylab="PD28 Arcsine Trans. Tot. Rel. Abund.",
        xlab="Treatment")
plotNormalHistogram(PD28_shared_otus$PD28_RA_logit)
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x)
# normal probability plot 
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

```{r}
################# PD28 Count #################
model = lmer(PD28_shared_otus$Count ~ TREATMENT + (1|MOM), data=PD28_shared_otus)
summary(model)
sstable <- anova(model)

# Creating an html ANOVA table 
options(knitr.kable.NA = '') # this will hide missing values in the kable table
kable(sstable, digits = 4, caption = "Count ANOVA table for PD28") %>%
  kable_styling(full_width = FALSE)

# Create 2x2 plot environment
par(mfrow = c(2, 2))
# plot data
boxplot(Count ~ TREATMENT, data = PD28_shared_otus,
        ylab="PD28 Count",
        xlab="Treatment")
# plot residules histogram
x = (residuals(model))
plotNormalHistogram(x)
# plot Q plot
qqnorm(residuals(model), ylab="Sample Quantiles for residuals")
qqline(residuals(model), col="red")
# simple scatterplot between residuals and predicted values. It should look more or less random.
plot(fitted(model), residuals(model))
```

### Plot Shared OTUs
```{r}
# Set plotting theme & colors
theme_set(theme_bw())
cols = c("Control" = "black", "MSEW" = "red")

# plot PD10 Count
pd10_count <- ggplot(PD10_shared_otus, aes(x=TREATMENT, y=Count)) +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 1.5)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "No. of shared\nASVs") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=14),axis.text.x=element_text(size=12)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
pd10_count 

# plot PD10 RA
pd10_RA <- ggplot(PD10_shared_otus, aes(x=TREATMENT, y=Tot.Rel.Abund.)) +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 1.5)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "Total relative abundance\nof shared ASVs") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=14),axis.text.x=element_text(size=12)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
pd10_RA

# plot PD10 Atch dist
pd10_atch <- ggplot(PD10_shared_otus, aes(x=TREATMENT, y=Distance)) +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 1.5)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "Aitchison\nDistance") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=14),axis.text.x=element_text(size=12)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
pd10_atch

# Set plotting theme & colors
theme_set(theme_bw())
cols = c("Control" = "black", "MSEW" = "red")

# plot PD28 Count
pd28_count <- ggplot(PD28_shared_otus, aes(x=TREATMENT, y=Count)) +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 1.5)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "No. of shared\nASVs") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=14),axis.text.x=element_text(size=12)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
pd28_count 

# plot PD28 RA
pd28_RA <- ggplot(PD28_shared_otus, aes(x=TREATMENT, y=Tot.Rel.Abund.)) +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 1.5)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "Total relative abundance\nof shared ASVs") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=14),axis.text.x=element_text(size=12)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
pd28_RA

# plot PD28 Atc Dist
pd28_atch <- ggplot(PD28_shared_otus, aes(x=TREATMENT, y=Distance)) +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour = TREATMENT), 
              position=position_jitter(width=0.1, height=0), size = 1.5)+
  scale_colour_manual(values = cols) +
  labs(x = " ",y = "Aitchison\nDistance") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size=14),axis.text.x=element_text(size=12)) +
  theme(legend.position="none") +
  theme(axis.title.y=element_text(colour="black",margin = margin(t = 0, r = 10, b = 0, l = 0)))
pd28_atch

################# Figure all panels #################
theme_set(theme_bw())
shared_outs_p <- plot_grid(pd10_count, pd28_count, 
                           pd10_RA, pd28_RA, 
                           pd10_atch, pd28_atch, 
                 ncol = 2, align = "hv")
shared_outs_p 
```

